---
title: "EDA of PFE"
---

```{r ,echo=FALSE, message=FALSE, warning=FALSE}
library(flipbookr)
library(tidyverse)
library(ggplot2)
library(forecast)
library(astsa) 
library(xts)
library(tseries)
library(fpp2)
library(fma)
library(lubridate)
library(tidyverse)
library(TSstudio)
library(quantmod)
library(tidyquant)
library(plotly)
require(gridExtra)

options("getSymbols.warning4.0"=FALSE)
options("getSymbols.yahoo.warning"=FALSE)

tickers = c("PFE","UNH","JNJ", "AZN", "SYK", "CVS","LLY")
for (i in tickers){
  getSymbols(i,
             from = "2010-1-1",
             to = "2023-1-29")}

x <- list(
  title = "date"
)
y <- list(
  title = "value"
)

stock <- data.frame(PFE$PFE.Adjusted,
                    UNH$UNH.Adjusted,
                    AZN$AZN.Adjusted,
                    SYK$SYK.Adjusted,
                    CVS$CVS.Adjusted,
                    JNJ$JNJ.Adjusted,
                    LLY$LLY.Adjusted)


stock <- data.frame(stock,rownames(stock))
colnames(stock) <- append(tickers,'Dates')

stock$date<-as.Date(stock$Dates,"%Y-%m-%d")
#write.csv(stock, 'data/stock.csv', row.names = FALSE)

stock_df <- read.csv('data/stock.csv')
stock_df$Dates <- as.Date(stock_df$Dates)
```

Refer to EDA-UNH for more detailed description for each plot.

### Basic Time Series Plot

```{r}
# candlestick plot

PFE_df <- as.data.frame(PFE)
PFE_df$Dates <- as.Date(rownames(PFE_df))

fig_PFE <- PFE_df %>% plot_ly(x = ~Dates, type="candlestick",
          open = ~PFE.Open, close = ~PFE.Close,
          high = ~PFE.High, low = ~PFE.Low) 
fig_PFE <- fig_PFE %>% 
  layout(title = "Basic Candlestick Chart for Pfizer")

fig_PFE
```

### Lag plot

```{r}
PFE_ts <- ts(stock_df$PFE, start = c(2010,1),end = c(2023,1),
             frequency = 251)
ts_lags(PFE_ts)
```

### Decomposed times series

```{r}
decompose_PFE <- decompose(PFE_ts,'multiplicative')
autoplot(decompose_PFE)
```

### Autocorrelation in Time Series

```{r}
PFE_acf <- ggAcf(PFE_ts,100)+ggtitle("ACF Plot for PFE")
PFE_pacf <- ggPacf(PFE_ts)+ggtitle("PACF Plot for PFE")

grid.arrange(PFE_acf, PFE_pacf,nrow=2)
```

### Augmented Dickey-Fuller Test

```{r}
tseries::adf.test(PFE_ts)
```

### Detrending

```{r}
fit = lm(PFE_ts~time(PFE_ts), na.action=NULL) 

y= PFE_ts
x=time(PFE_ts)
DD<-data.frame(x,y)
ggp <- ggplot(DD, aes(x, y)) +           
  geom_line()

ggp <- ggp +                                     
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +ggtitle("PFE Stock Price")+ylab("Price")

plot1<-autoplot(resid(fit), main="detrended") 
plot2<-autoplot(diff(PFE_ts), main="first difference") 


grid.arrange(ggp, plot1, plot2,nrow=3)
```

### Moving Average Smoothing

Smoothing methods are a family of forecasting methods that average values over multiple periods in order to reduce the noise and uncover patterns in the data. It is useful as a data preparation technique as it can reduce the random variation in the observations and better expose the structure of the underlying causal processes. We call this an m-MA, meaning a moving average of order m.

```{r,message=FALSE,warning=FALSE}
MA_7 <- autoplot(PFE_ts, series="Data") +
        autolayer(ma(PFE_ts,7), series="7-MA") +
        xlab("Year") + ylab("Adjusted Closing Price") +
        ggtitle("PFE Stock Price Trend in (7-days Moving Average)") +
        scale_colour_manual(values=c("PFE_ts"="grey50","7-MA"="red"),
                            breaks=c("PFE_ts","7-MA"))

MA_30 <- autoplot(PFE_ts, series="Data") +
        autolayer(ma(PFE_ts,30), series="30-MA") +
        xlab("Year") + ylab("Adjusted Closing Price") +
        ggtitle("PFE Stock Price Trend in (30-days Moving Average)") +
        scale_colour_manual(values=c("PFE_ts"="grey50","30-MA"="red"),
                            breaks=c("PFE_ts","30-MA"))

MA_251 <- autoplot(PFE_ts, series="Data") +
        autolayer(ma(PFE_ts,251), series="251-MA") +
        xlab("Year") + ylab("Adjusted Closing Price") +
        ggtitle("PFE Stock Price Trend in (251-days Moving Average)") +
        scale_colour_manual(values=c("PFE_ts"="grey50","251-MA"="red"),
                            breaks=c("PFE_ts","251-MA"))

grid.arrange(MA_7, MA_30, MA_251, ncol=1)
```

The graph above shows the moving average of 7 days, 30 days and 251 days. 251 days was choose because there are around 251 days of stock price data per year. According to the plots, it can be observed that When MA is very large(MA=251), some parts of smoothing line(red) do not fit the real stock price line. While When MA is small(MA=7), the smoothing line(red) fits the real price line. MA-30 greatly fits the real price line. Therefore, MA-30 might be a good parameter for smoothing.
